Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-16 paper.tex
Edit evaluation

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage{newunicodechar}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{hyperref}
\usepackage{csquotes}
\usepackage[noend]{algorithm2e}

@@

\newunicodechar{₀}{\ensuremath{_{0}}}
\newunicodechar{₁}{\ensuremath{_{1}}}
\newunicodechar{₂}{\ensuremath{_{2}}}
\newunicodechar{₃}{\ensuremath{_{3}}}
\newunicodechar{₄}{\ensuremath{_{4}}}
\newunicodechar{₅}{\ensuremath{_{5}}}
\newunicodechar{₆}{\ensuremath{_{6}}}
\newunicodechar{₇}{\ensuremath{_{7}}}
\newunicodechar{₈}{\ensuremath{_{8}}}
\newunicodechar{₉}{\ensuremath{_{9}}}
\newunicodechar{ₙ}{\ensuremath{_{n}}}
\newunicodechar{ₘ}{\ensuremath{_{m}}}
\newunicodechar{ᵢ}{\ensuremath{_{i}}}
\newunicodechar{ⱼ}{\ensuremath{_{j}}}
\newunicodechar{ₗ}{\ensuremath{_{l}}}
\newunicodechar{ₖ}{\ensuremath{_{k}}}
\newunicodechar{α}{\ensuremath{\alpha}}
\newunicodechar{σ}{\ensuremath{\sigma}}
\newunicodechar{τ}{\ensuremath{\tau}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{μ}{\ensuremath{\mu}}
\newunicodechar{η}{\ensuremath{\eta}}
\newunicodechar{Γ}{\ensuremath{\Gamma}}
\newunicodechar{Δ}{\ensuremath{\Delta}}
\newunicodechar{Φ}{\ensuremath{\Phi}}
\newunicodechar{Ψ}{\ensuremath{\Psi}}
\newunicodechar{Ω}{\ensuremath{\Omega}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∉}{\ensuremath{\notin}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≢}{\ensuremath{\nequiv}}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{≥}{\ensuremath{\geq}}
\newunicodechar{≔}{\ensuremath{\coloneq}}
\newunicodechar{∧}{\ensuremath{\land}}
\newunicodechar{∨}{\ensuremath{\lor}}
\newunicodechar{∪}{\ensuremath{\cup}}
\newunicodechar{∩}{\ensuremath{\cap}}
\newunicodechar{⊆}{\ensuremath{\subseteq}}
\newunicodechar{⋃}{\ensuremath{\bigcup}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{↔}{\ensuremath{\leftrightarrow}}
\newunicodechar{↦}{\ensuremath{\mapsto}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{∅}{\ensuremath{\emptyset}}
\newunicodechar{∷}{\ensuremath{\mathbin{::}}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{⟨}{\ensuremath{\langle}}
\newunicodechar{⟩}{\ensuremath{\rangle}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{×}{\ensuremath{\times}}

@@

\section{Incremental Forward Reasoning}

@@

Comparing the stateless and stateful approach, we can draw the following conclusions about their relative performance:

\begin{itemize}
    \item The stateful algorithm maybe be expected to perform poorly when the context contains many hypotheses that almost complete rules.
    \item The stateless algorithm should perform decently well on average and especially when only the last premises are absent, so the rules get never selected by the stateless index.
\end{itemize}

These tests were performed on a Macbook Pro with the Apple M2 Pro chip equipped with 32 Go of RAM.
The tests showcased in this section are available here \footnote{\xcom{TODO : Link Github repo with tests.}}
\subsection{Results}
\xcom{TODO: Add a comment on the effect it had on Mathlib.}

We present here a number of results obtained by running synthetic tests.
By synthetic we mean that these tests we're constructed solely for the purpose of demonstrating the different complexity behavior of both algorithms by scaling some parameter (e.g.\ the number of rules).
The results of the tests are an average of multiple runs.

\paragraph{Transitivity.}

\begin{figure}
    \label{fig:trans}
    % Transitivity without precompilation.
    % 
    \begin{tikzpicture}
        \begin{axis}[
            title={Transitivity rule},
            xlabel={Number of hypotheses},
            ylabel={Time in ms},
            xmin=1, xmax=16,
            ymin=2^3, ymax=2^13,
            xtick={1,2,4,8,16},
            ytick={2^3,2^5,2^7,2^9,2^11,2^13},
            xmode=log,
            ymode=log,
            log basis x=2,
            log basis y=2,
            legend pos=north west,
            ymajorgrids=true,
            grid style=dashed,
            legend image post style={mark=},
        ]

        \addplot[
            color=orange,
            mark=square,
            ]
            coordinates {
              (1, 7.663472) (2, 9.354750) (4, 15.768194) (8, 104.921791) (16, 1882.683111)
            };
            \addlegendentry{Stateless}
        \addplot[
            color=blue,
            mark=square,
            ]
            coordinates {
              (1, 10.308541) (2, 10.704638) (4, 16.274305) (8, 38.800833) (16, 179.697681) 
            };
            \addlegendentry{Stateful}
            \addplot[
              color=orange,
              mark=o,
              ]
              coordinates {
                (1, 10.591138) (2, 23.881333) (4, 142.013389) (8, 1676.148375) (16, 24327.359417)
              };
          \addplot[
              color=blue,
              mark=o,
              ]
              coordinates {
                (1, 13.974694) (2, 26.216833) (4, 68.953444) (8, 245.601028) (16, 1028.670680) 
              };
        \end{axis}
        \end{tikzpicture}
    \end{figure}

As a first example, which we believe highlights well the advantages of the stateful approach, we consider a scaled up version of a transitivity problem.
As variables we choose a custom version of natural numbers.
The reason being that Lean has special optimization procedures for its integer and thus makes evaluation quite hard on problems that use them; the unification is too quickly resolved, making algorithmic complexity difficultly distinguishable from noise.
We aim to emulate a more general problem where the objects unified do not benefit of some special support.

Given some predicate $P(x, y)$, we register the rule $r : P(x, y) \to P(y, z) \to P(x, z)$ to both algorithms.
Then, in Figure \ref{fig:trans}, we compare them on a context containing $n$ hypotheses of the form $(h_1 : P(a, n + 1)), \dots, (h_n : P(a + n - 1, a + n))$.
Saturating this goal adds a total of $n(n-1)/2$ hypotheses to the context.

The results in Figure \ref{fig:trans} have nothing surprising : The stateless algorithm needs to go over the whole context, which contains initially $n$ hypotheses and compares them with these same $n$ hypotheses.
This is repeated approximately for each of the $n(n-1)/2$ added hypotheses giving the expected complexity of $O(n^4)$.

The stateful algorithm considers each of the problem's hypothesis, both the initial and added ones, only once.
As the hypotheses get matched to both slots, it then compares it with at most $n$ hypotheses for each slot.
Indeed, by the end of the saturation procedure, for a value $k$ fixed, the first slot $\eta_y(1,k)$, contains at most $n$ hypotheses of the form $h : P(m,k)$ and symmetrically for the second slot.
This leaves an expected complexity of $O(n^2)$.

There are of course obvious inefficiencies, even with the stateful algorithm.
For this specific kind of problem, the hypotheses need only be used once and this leads to much repeated work.
For these scenarios, users can register the rule as a \textit{destruct} rule, eliminating hypotheses after their first use.
In this case, the complexity is linear in $n$.
For more edge case examples like this one, consult the project's repository.

\paragraph{Independent rules.}

\begin{figure}
    %Indep without precompilation
        \begin{tikzpicture}
            \begin{axis}[
                title={Independance Test},
                xlabel={Number of Rules},
                ylabel={Time in ms},
                xmin=1, xmax=32,
                ymin=2^3, ymax=1700,
                ymode=log,xmode=log,
                log basis x=2,
                log basis y=2,
                legend pos=north west,
                ymajorgrids=true,
                grid style=dashed,
                legend image post style={mark=},
            ]

            \addplot[
              color=orange,
              mark=square,
              ]
              coordinates {
                (1, 10.399958) (2, 16.426541) (4, 43.461403) (8, 205.595541) (16, 1238.090889) (32, 8927.857819) 
              };
              \addlegendentry{Stateless}
            \addplot[
                color=blue,
                mark=square,
                ]
                coordinates {
                  (1, 11.231875) (2, 12.879500) (4, 16.851930) (8, 24.666485) (16, 42.054055) (32, 76.369778) 
                };
                \addlegendentry{Stateful}
                \addplot[
              color=orange,
              mark=o,
              ]
              coordinates {
                (1, 19.405222) (2, 33.487472) (4, 80.107597) (8, 284.954791) (16, 1536.089624) (32, 10721.729861)
              };
            \addplot[
                color=blue,
                mark=o,
                ]
                coordinates {
                  (1, 25.892069) (2, 40.031930) (4, 69.446902) (8, 130.640986) (16, 260.653917) (32, 481.827902)
                };
            \end{axis}
            \end{tikzpicture}
        \end{figure}

In the last example, part of the stateful algorithm's success stems from the fact that it could reuse a lot of its previous work.
An obvious question is then how well does the stateful algorithm work when there is no work to be reused.

@@

\paragraph{Depth of considered rules.}

\begin{figure}
    \label{fig:erase}
%erase without precompilation
    \begin{tikzpicture}
        \begin{axis}[
            title={Depth Test},
            xlabel={Depth},
            ylabel={Time in ms},
            xmin=1, xmax=6,
            ymin=6, ymax=20,
            xtick={1,2,3,4,5,6},
            ytick={10,15},
            legend pos=north west,
            ymajorgrids=true,
            grid style=dashed,
            legend image post style={mark=},
        ]
        \addplot[
            color=orange,
            mark=square,
            ]
            coordinates {
              (1, 6.407973) (2, 9.831676) (3, 10.301718) (4, 10.626144) (5, 11.261754) (6, 11.795240) 
            };
            \addlegendentry{Stateless}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 15.835558) (2, 16.961994) (3, 16.424259) (4, 17.605449) (5, 15.691536) (6, 14.982820) 
          };
          \addlegendentry{Stateful}
        \end{axis}
        \end{tikzpicture}
    \end{figure}

In Fig.~\ref{fig:erase}, the step is set up as follows: given some rule $r$ with $6$ \xcom{motivate 6} premises, we run the procedure in a context with 5 of the 6 premises present.
The numbers on the $x$ axis represent which hypothesis is missing in the order of consideration by the algorithm.
In a sense, this is how \textit{deep} the algorithm investigate before it stops.
At $0$, the rule are not selected by the index so the procedure stops.
The stateless algorithm gets progressively slower the further the removed premises is located.

@@

\paragraph{SATCHMO.}
The Prolog-based first-order theorem prover SATCHMO (TODO ref) and its many descendants (TODO refs) uses a tableau calculus (TODO ref) that can be viewed as combined forward and backward reasoning.
Given a set of implications $A₁ ∧ \dots ∧ Aₙ → B₁ ∨ \dots ∨ Bₘ$, SATCHMO attempts to create a model that satisfies all implications.
Its search procedure is a special case of our technique: the implications are used as forward rules and a single backward rule splits any goal with target $B₁ ∨ \dots ∨ Bₘ$ into goals with targets $B₁, \dots, Bₘ$.
This is a simpler setting than ours since the backward rule does not change the facts available to the forward rules.
Still, our technique has the advantage that at each branching point we remember the partial matches induced by the currently known facts for each implication, rather than matching the implication against the database multiple times.

>>>>

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage{newunicodechar}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{hyperref}
\usepackage{csquotes}
\usepackage[noend]{algorithm2e}
\usepackage{siunitx}

@@

\newunicodechar{₀}{\ensuremath{_{0}}}
\newunicodechar{₁}{\ensuremath{_{1}}}
\newunicodechar{₂}{\ensuremath{_{2}}}
\newunicodechar{₃}{\ensuremath{_{3}}}
\newunicodechar{₄}{\ensuremath{_{4}}}
\newunicodechar{₅}{\ensuremath{_{5}}}
\newunicodechar{₆}{\ensuremath{_{6}}}
\newunicodechar{₇}{\ensuremath{_{7}}}
\newunicodechar{₈}{\ensuremath{_{8}}}
\newunicodechar{₉}{\ensuremath{_{9}}}
\newunicodechar{ₙ}{\ensuremath{_{n}}}
\newunicodechar{ₘ}{\ensuremath{_{m}}}
\newunicodechar{ᵢ}{\ensuremath{_{i}}}
\newunicodechar{ⱼ}{\ensuremath{_{j}}}
\newunicodechar{ₗ}{\ensuremath{_{l}}}
\newunicodechar{ₖ}{\ensuremath{_{k}}}
\newunicodechar{α}{\ensuremath{\alpha}}
\newunicodechar{σ}{\ensuremath{\sigma}}
\newunicodechar{τ}{\ensuremath{\tau}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{μ}{\ensuremath{\mu}}
\newunicodechar{η}{\ensuremath{\eta}}
\newunicodechar{Γ}{\ensuremath{\Gamma}}
\newunicodechar{Δ}{\ensuremath{\Delta}}
\newunicodechar{Φ}{\ensuremath{\Phi}}
\newunicodechar{Ψ}{\ensuremath{\Psi}}
\newunicodechar{Ω}{\ensuremath{\Omega}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∉}{\ensuremath{\notin}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≢}{\ensuremath{\nequiv}}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{≥}{\ensuremath{\geq}}
\newunicodechar{≺}{\ensuremath{\prec}}
\newunicodechar{≔}{\ensuremath{\coloneq}}
\newunicodechar{∧}{\ensuremath{\land}}
\newunicodechar{∨}{\ensuremath{\lor}}
\newunicodechar{∪}{\ensuremath{\cup}}
\newunicodechar{∩}{\ensuremath{\cap}}
\newunicodechar{⊆}{\ensuremath{\subseteq}}
\newunicodechar{⋃}{\ensuremath{\bigcup}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{↔}{\ensuremath{\leftrightarrow}}
\newunicodechar{↦}{\ensuremath{\mapsto}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{∅}{\ensuremath{\emptyset}}
\newunicodechar{∷}{\ensuremath{\mathbin{::}}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{⟨}{\ensuremath{\langle}}
\newunicodechar{⟩}{\ensuremath{\rangle}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{×}{\ensuremath{\times}}

@@

\section{Incremental Forward Reasoning}%
\label{sec:incremental}

@@

There are unfortunately no standard benchmarks for Aesop-like tactics, and since Aesop is not expected to solve most goals without an extensive, manually curated set of rules, benchmarks for push-button theorem provers such as TPTP (TODO ref) are not very meaningful.
At the same time, forward rules are currently little used by Aesop's main client, the Mathlib mathematics library (TODO ref).
We therefore evaluate our technique against the previous, naive Aesop implementation on several synthetic benchmarks that showcase the performance characteristics of incremental forward reasoning.

For each benchmark, we set up certain forward rules and a goal with certain hypotheses.
We then run the \texttt{saturate} tactic, which uses either naive or incremental forward reasoning to exhaustively apply the forward rules.
Performing only forward reasoning is the best-case scenario for the incremental algorithm since previous results can be fully reused.
When forward reasoning is combined with other rules, we would expect smaller gains, depending on how much reuse the non-forward rules allow.

Benchmark code and data are available in our supplement (TODO ref).
The results shown below were obtain on a Macbook Pro with an Apple~M2 Pro processor and \SI{32}{GB} of RAM, averaging over \textcolor{red}{N} (TODO) runs per benchmark.

\paragraph{Transitivity.}
Our first benchmark is a scaled-up variant of the transitivity example from Sec.~\ref{sec:incremental}.
Given some relation $≺$ on natural numbers, we register the lemma $\All{x~y}{x ≺ y → y ≺ z → x ≺ z}$ as the sole forward rule and set up a goal with hypotheses $h₁ : a ≺ a + 1, \dots, hₙ : a + n - 1 ≺ a + n$ (TODO what is $a$?).
Saturating this goal adds $n(n-1)/2$ hypotheses to the context.

Instead of the natural number type of Lean's standard library, which has special support in the compiler and kernel, we use a custom type of Peano naturals.
Unifying two concrete Peano natural numbers $n$ and $m$ involves comparing $\min(n, m)$ constructors.
We can therefore use moderately large numbers to simulate the bigger unification problems that Aesop encounters, for example, in Mathlib's category theory development.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={Number of hypotheses},
        ylabel={Time in ms},
        xmin=1, xmax=16,
        ymin=2^3, ymax=2^13,
        xtick={1,2,4,8,16},
        ytick={2^3,2^5,2^7,2^9,2^11,2^13},
        xmode=log,
        ymode=log,
        log basis x=2,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 7.663472) (2, 9.354750) (4, 15.768194) (8, 104.921791) (16, 1882.683111)
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 10.308541) (2, 10.704638) (4, 16.274305) (8, 38.800833) (16, 179.697681)
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 10.591138) (2, 23.881333) (4, 142.013389) (8, 1676.148375) (16, 24327.359417)
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 13.974694) (2, 26.216833) (4, 68.953444) (8, 245.601028) (16, 1028.670680)
          };
      \end{axis}
    \end{tikzpicture}
  \end{center}
  \caption{Results of the transitivity benchmark. TODO detailed description}%
  \label{fig:trans}
\end{figure}

The results in Fig.~\ref{fig:trans} are unsurprising.
The naive algorithm goes over the context, containing initially $n$ hypotheses, and compares all hypotheses pairwise.
This is repeated approximately for each of the $n(n-1)/2$ added hypotheses, so the overall complexity is $O(n^4)$.
The incremental algorithm considers each hypothesis only once.
As the hypotheses get matched to both slots, it then compares each hypothesis with at most $n$ other hypotheses.
This yields a complexity of $O(n^2)$.

Either algorithm generates all possible transitivity chains.
If only the longest such chain, from $a$ to $a + n$, is needed, the transitivity rule can be registered as a \emph{destruct} rule, a special sort of forward rule that, whenever it is applied to certain hypotheses, removes the hypotheses from the context.
This reduces the complexity of the incremental algorithm to $O(n)$.

\paragraph{Independent Rules.}
In the last example, part of the stateful algorithm's success stems from the fact that it could reuse a lot of its previous work.
An obvious question is then how well does the stateful algorithm work when there is no work to be reused.

@@

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={Number of Rules},
        ylabel={Time in ms},
        xmin=1, xmax=32,
        ymin=2^3, ymax=1700,
        ymode=log,xmode=log,
        log basis x=2,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]

        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 10.399958) (2, 16.426541) (4, 43.461403) (8, 205.595541) (16, 1238.090889) (32, 8927.857819)
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 11.231875) (2, 12.879500) (4, 16.851930) (8, 24.666485) (16, 42.054055) (32, 76.369778)
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 19.405222) (2, 33.487472) (4, 80.107597) (8, 284.954791) (16, 1536.089624) (32, 10721.729861)
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 25.892069) (2, 40.031930) (4, 69.446902) (8, 130.640986) (16, 260.653917) (32, 481.827902)
          };
      \end{axis}
    \end{tikzpicture}
  \end{center}
  \caption{Results of the independence benchmark. TODO detailed description}%
  \label{fig:indep}
\end{figure}

\paragraph{Rules of Varying Depth.}
In Fig.~\ref{fig:erase}, the step is set up as follows: given some rule $r$ with $6$ \xcom{motivate 6} premises, we run the procedure in a context with 5 of the 6 premises present.
The numbers on the $x$ axis represent which hypothesis is missing in the order of consideration by the algorithm.
In a sense, this is how \textit{deep} the algorithm investigate before it stops.
At $0$, the rule are not selected by the index so the procedure stops.
The stateless algorithm gets progressively slower the further the removed premises is located.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        ylabel={Time in ms},
        xmin=1, xmax=6,
        ymin=6, ymax=20,
        xtick={1,2,3,4,5,6},
        ytick={10,15},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 6.407973) (2, 9.831676) (3, 10.301718) (4, 10.626144) (5, 11.261754) (6, 11.795240)
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 15.835558) (2, 16.961994) (3, 16.424259) (4, 17.605449) (5, 15.691536) (6, 14.982820)
          };
        \addlegendentry{incremental}
      \end{axis}
    \end{tikzpicture}
  \end{center}
  \caption{Results of the depth benchmark. TODO detailed description}%
  \label{fig:erase}
\end{figure}

@@

\paragraph{SATCHMO.}
The Prolog-based first-order theorem prover SATCHMO (TODO ref) and its many descendants (TODO refs) use a tableau calculus (TODO ref) that can be viewed as combined forward and backward reasoning.
Given a set of implications $A₁ ∧ \dots ∧ Aₙ → B₁ ∨ \dots ∨ Bₘ$, SATCHMO attempts to create a model that satisfies all implications.
Its search procedure is a special case of our technique: the implications are used as forward rules and a single backward rule splits any goal with target $B₁ ∨ \dots ∨ Bₘ$ into goals with targets $B₁, \dots, Bₘ$.
This is a simpler setting than ours since the backward rule does not change the facts available to the forward rules.
Still, our technique has the advantage that at each branching point we remember the partial matches induced by the currently known facts for each implication, rather than matching the implication against the database multiple times.
